---
title: "Choose time slots for a talk"
author: "Tomás E. Tecce"
output: html_notebook
---

## The problem

I need to give an introductory talk to a class of students who will be
taking an online course. I have already selected a number of possible time
slots for the talks, based on my own availability. The students are
located in different countries and time zones, and I'm willing to repeat
my talk at least once in order to accomodate this.

I've asked the students to complete a poll, indicating which of the time
slots they prefer; they can choose more than one option. What I would like
to do is to use that information to choose two time slots in order to
maximize the number of students who can attend the talk.


## The data

The poll data was collected using [Doodle](http://doodle.com). People were
asked to choose among 30 possible time slots, and multiple choices were
allowed. The results are collected in a CSV file which contains one row
per person, and the columns correspond to the different possible time
slots (labeled from S01 to S30).

```{r}
# Read the data file
df <- read.csv("data.csv", na.strings="", stringsAsFactors=FALSE)

# Convert the data from "OK" and NA to 0 and 1, then find the total number
# of people who chose each of the different time slots
df[is.na(df)] <- 0
df[df == "OK"] <- 1
df[,2:31] <- apply(df[,2:31], 2, as.numeric)

ts_sums <- colSums(df)
```

The maximum number of people choosing the same time slot is
`r max(ts_sums[2:31])`, which is much less than the total number of
students (`r dim(df)[1]`). Thus I cannot get away with giving just one
talk if I want at least 50 per cent of the students to attend the course
introduction.

```{r}
library(ggplot2)

fig <- ggplot(data=data.frame(timeslots=colnames(df)[2:31],
                              counts=ts_sums[2:31]),
              aes(x=timeslots, y=counts)) + geom_bar(stat="identity", fill="steel blue")
print(fig)
```

## The method

To find two timeslots that maximize the number of distinct students that
attend the introduction, I use the following method.

First, I build new columns, one for each possible permutation of time
slots (slot #1 and slot #2, slot #1 and slot #3, etc). Each row contains 1
if the student can attend either of the slots, and 0 otherwise. 

```{r}
library(dplyr)

new_names <- sapply(combn(names(df[,2:31]), 2, simplify = FALSE),
                    function(x) { paste(x, collapse="") })

for (i in 1:length(new_names)) {
  df <- mutate(df, newcol=as.numeric(df[substr(new_names[i], 1, 3)] |
                                       df[substr(new_names[i], 4, 6)]))
  colnames(df)[colnames(df)=="newcol"] <- new_names[i]
}
```

Then we sum the columns to find which time slot combination has the most
attendee count.

```{r}
ts_sums <- colSums(df[,32:dim(df)[2]])
ts_sums[order(-ts_sums)][1:10]
```

Thus, we find that by choosing slots #3 and #30 we could potentially have
63 students attending (54 per cent of the total).

We can repeat the exercise for 3 time slots:

```{r}
new_names <- sapply(combn(names(df[,2:31]), 3, simplify = FALSE),
                    function(x) { paste(x, collapse="") })
enddim <- dim(df)[2]

for (i in 1:length(new_names)) {
  df <- mutate(df, newcol=as.numeric(df[substr(new_names[i], 1, 3)] |
                                       df[substr(new_names[i], 4, 6)] |
                                       df[substr(new_names[i], 7, 9)]))
  colnames(df)[colnames(df)=="newcol"] <- new_names[i]
}

ts_sums <- colSums(df[,enddim+1:dim(df)[2]])
ts_sums[order(-ts_sums)][1:10]
```

